---
title: "Hypothesis Testing Problem using R"
author: "Eriko Funasato"
date: "October 31, 2017"
output: github_document
---
#Introduction
  In elementary of statistic, we have learnt about hypothesis testing.A statistical hypothesis test is a method of statistical inference. Commonly, two statistical data sets are compared, or a data set obtained by sampling is compared against a synthetic data set from an idealized model([Wikipedia](https://en.wikipedia.org/wiki/Statistical_hypothesis_testing)).We compare one alternative statement against a null hypothesis that is supposed to be true. In hypothesis testing, we determined if alternative claim is true by calculating whether the null hypothesis should be rejected or not. 
<br><br>
  In this process, we could introduce terms such as types of errors, significant level, p-value, and some graphs to make us easier to analyze the dataset. In this post, I am going to apply R into hypothesis testing problem by some example using the  dataset "earn-of-collegemajors-all-ages"([Github](https://github.com/fivethirtyeight/data/blob/master/college-majors/all-ages.csv)).
--import the dataset and packages to be used 
```{r}
dat <- read.csv("C:/Users/eriko/stat133/stat133-hws-fall17/post01/data/earn-of-collegemajors-all-ages.csv",stringsAsFactors=FALSE)
library(dplyr)
library(ggplot2)
library(BSDA)
setwd("/Users/eriko/stat133/stat133-hws-fall17/post01")
```
<br><br><br>
#Graphing and Analyze
* Before we start doing a hypothesis testing problem, let's organize the data best fit to be worked in R studio, and do some analysis and graphing process to know more about the dataset. 
<br><br>
* check the structure of the dataset.
```{r}
str(dat)
```
<br><br>
* Rename the columns
The name of columns in the dataset is too long and complicated, it is better to replace them by shorter abbreviation. 
```{r}
#replace the major_category's name and column name
colnames(dat) <- c("major_code","major","major_cat","total","employed","employed_fulltime","unemployed","unemployed_rate","median","q1","q3")
```
<br><br>
* Add column of employment rate
```{r}
dat <- mutate(dat,employed_rate = dat$employed/dat$total)
```
<br><br>
* Picking data of "Computers & Mathematics"and "engineering"
```{r}
#Create new dataframes of category of "Computers & Mathematics"and "engineering"
cs_math <- filter(dat,dat$major_cat=="Computers & Mathematics")
engin <- filter(dat,dat$major_cat=="Engineering")
#store to data file
write.csv(cs_math,file = "C:/Users/eriko/stat133/stat133-hws-fall17/post01/data/cs_math.csv")
write.csv(engin,file = "C:/Users/eriko/stat133/stat133-hws-fall17/post01/data/engin.csv")
```
<br><br>
* summary of data of "Computers & Mathematics"and "engineering"
```{r}
#store the result in output file
sink(file = "/Users/eriko/stat133/stat133-hws-fall17/post01/output/cs_math-summary.txt")
summary(cs_math)
sink()
sink(file = "/Users/eriko/stat133/stat133-hws-fall17/post01/output/summary-engin.txt")
summary(engin)
sink()
```
<br><br>
* Plot and store employment rate of each dataframe
```{r}
ggplot(cs_math)+geom_col(aes(x=major,y=employed_rate),color=NA,fill="pink")+labs(x="Majors of computer science and math",y="Employment rate")
```

```{r}
ggplot(engin)+geom_col(aes(x=major,y=employed_rate),color=NA,fill="blue")+labs(x="Majors of college of engineering",y="Employment rate")
```

```{r}
#save
pdf(file = "C:/Users/eriko/stat133/stat133-hws-fall17/post01/image/csmath-ggplot.pdf")
ggplot(cs_math)+geom_col(aes(x=major,y=employed_rate),color=NA,fill="pink")+labs(x="Majors of computer science and math",y="Employment rate")
dev.off()
```
```{r}
#save
pdf(file = "C:/Users/eriko/stat133/stat133-hws-fall17/post01/image/engin-ggplot.pdf")
ggplot(engin)+geom_col(aes(x=major,y=employed_rate),color=NA,fill="blue")+labs(x="Majors of college of engineering",y="Employment rate")
dev.off()
```
<br><br><br><br>
# Hypothesis
* To build up my own hypothesis testing problem, I would like to compare which one of "Computer science & math" and "Engineering" has higher income salary, since I was wondering for a long time if which of the field is paid more.
<br><br>
* But to compare the income salary by hypothesis test, I need the mean and standard deviation of the datasets.
<br><br>  
* To get the mean and sd, here is the formula ([Wan et al.(2014)](https://bmcmedresmethodol.biomedcentral.com/articles/10.1186/1471-2288-14-135)),

<br>$\bar{x}=\frac{q1+m+q3}{3}$
<br><br>$S=\frac{q3???q1}{1.35}$
<br><br>where q1 is first quartile, m is the median, q3 is the third quartile
+ Add the mean and standard deviation to each datasets
```{r}
cs_math <- mutate(cs_math,mean=(q3+median+q1)/3,std=(q3-q1)/1.35)
engin <- mutate(engin,mean=(q3+median+q1)/3,std=(q3-q1)/1.35)
```
+ Calculate the mean and the std of whole datasets
```{r}
Mean_cs_math <- sum(cs_math$mean * cs_math$total) / sum(cs_math$total)
Mean_engin <- sum(engin$mean * engin$total) / sum(engin$total)
std_cs_math <- mean(cs_math$std)
std_engin <- mean(engin$std)
```

####Hypothesis: Computer science & math earn more money than the engineering major. 

1. Claim<br> 
H_0 (null): mean(cs_math income)-mean(engineering income) = 0<br/>
H_1 (alternative claim): mean(cs_math income)-mean(engineering income) > 0
2. since the number of both independent sample are huge enough, suppose they are approximately normaly distributed. Standard deviation are known by calculation, we could use 2-sample Z-test.
3. Set the significant value as alpha=0.05, which means confidence level is 0.95
4. Program calculation
+ Calculate using z.test([RDocumentation](https://www.rdocumentation.org/packages/BSDA/versions/1.2.0/topics/z.test))
```{r}
z.test(cs_math$mean,engin$mean, alternative = "greater", mu = 0, sigma.x = std_cs_math,sigma.y = std_engin, conf.level = 0.95)
```
+ Calculate using t.test([Phil Spector](https://statistics.berkeley.edu/computing/r-t-tests))
```{r}
t.test(cs_math$mean,engin$mean,alternative = "greater",conf.level = 0.95)
```
5. Conclusion: fail to reject null
* For the result of calculation using z-test, p-value=0.7854 is greater than the alpha=0.05, therefore fail to reject the null hypothesis, we don't have enough confidence to say that people who major in "Computer science & math" earn more than "Engineering" major.
<br>
* For the result of calculation using t-test, p-value=0.9906 is greater than the alpha=0.05, therefore fail to reject the null hypothesis, we don't have enough confidence to say that people who major in "Computer science & math" earn more than "Engineering" major.
<br>Therefore, we cannot conclude which of the major("CS & Math" and "Engineering") can earn more money after graduate.
<br>
* The difference between the Z-test and t-test is, when the mean of the sample can be known, the standard deviation is unknown, then we use the **t-test** to estimate the population. If the standard deviation is known or calculated, then use **z-test** to do the evaluation.
<br><br>


# More about hypothesis test.
* Not only the problems like the example I give, there are still many kind of hypothesis tesing, such as:<br/>
  + one-sample z-test(left-sided,right-sided,both-sided)
  + one-sample t-test(left-sided,right-sided,both-sided)
  + two-proportion z test(inference of two proportion)
  + two-proportion t-test
  + two-sample t/z-test(inference of two mean)
  + chi-squre test(testinga claim about a standard deviation of population)
  + Goodness-of-fit(test if sample data with k categories is "good fit"to an assumed distribution)
  + One-way Analysis of Variance(ANOVA):test for equality of more than three sample data.
<br>
* And to do these in r, check ([Rtutorial](http://www.r-tutor.com/elementary-statistics/hypothesis-testing))<br/>
For the concepts of hypothesis test, see ([Pennstate](https://onlinecourses.science.psu.edu/statprogram/node/139)).
<br><br>

# Take Home Message!!
PLease try to make some function of t-test or z-test, for one-sample and two-sample.<br/>
<br>
And try to compare the employment rate and unemployment rate of the majors or the major categories you are interested in. 
<br><br><br><br>
**Thank you for reading my post!!!**

